Online Model-Based and Model-Free Reinforcement Learning in Accelerator Operation with Applications to FERMI FEL

### 2020-12-03 main.tex
- paper changes

		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],

@@

	\section{Experimental results from FERMI RL online tests}
	As discussed in the previous sections, several tests were performed on the FERMI FEL.
	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility. The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
	\subsection{MFRL tests}
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
		\caption{The training of the \emph{NAF2} on the FERMI FEL.}
		\label{fig:NAF_training}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
		\caption{The verification of the \emph{NAF2} on the FERMI FEL.}
		\label{fig:NAF_verification}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
		\caption{The verification of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:NAF_convergence}
	\end{figure}
	In total, four tests were carried out, two using a single network and two using the double network architecture. A motivation for the used algorithmic set-up is discussed in \cref{appendix:naf2} on a classical example from control theory. In both cases \emph{smoothing} to increase the noise robustness as discussed in \cref{appendix:naf2}) was applied.\\
	\Cref{fig:NAF_training} displays the results, averaged over the two tests. A training of 100 episodes was accomplished. In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps. The green line in the lower plot of \cref{fig:NAF_training} indicates the 95\% intensity reward threshold, to accept an episode to be successfully finished.
	In the verification of 50 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance. \\
	Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps. The blue curves show the Bellmann error, which is comparable in both cases. The state-value function, which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value. In the single-network case, convergence is reached ~1400 steps.
	\subsection{MBRL tests}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
		\caption{The training of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:AE-DYNA_observables}
	\end{figure}

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference clean up.
	\item Rethink the concept of an epoch.
	\item Conclusion and outlook\dots
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}

>>>>

		V^\policy(\bs_t) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],\label{eq:state-value-function}

@@

	\subsection{Comparison of the used algorithms}
	\Cref{tab:overview_algorithms} shows a comparison of the discussed algorithms. Details are provided in the appendix. All algorithms are suitable for many control tasks as common in accelerator operation. Their strength lies in the high sample efficiency, but they differ in representational power and noise robustness. \\
	In \cref{sec:Experimental results from FERMI RL online tests} the \emph{NAF2},
	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.
		\begin{table}[b]%The best place to locate the table environment is directly after its first reference in text
	\caption{\label{tab:overview_algorithms}%
		An overview of the algorithms discussed in this paper.
	}
	\begin{ruledtabular}
		\begin{tabular}{lllll}
			\textrm{Algorithm}&
			\textrm{Class}&
			\textrm{NR\footnote{Noise robustness}}&
			\textrm{RP\footnote{Representational power}}&
			\textrm{SE\footnote{Sample efficiency}}\\
			\colrule
			\emph{NAF} & model-free & low &low &high\\
			\emph{NAF2} & model-free & high &low& high\\
			\emph{ME-TRPO} & model-based & low& high& high\\
			\emph{AE-DYNA} & model-based & high& high& high\\
		\end{tabular}
	\end{ruledtabular}
\end{table}
	\section{Experimental results from FERMI RL online tests}\label{sec:Experimental results from FERMI RL online tests}
	As discussed in the previous sections, several tests were performed on the FERMI FEL.
	The main purpose was to test the newly implemented algorithms on a real system to evaluate its operational feasibility.\\ 
	Because of the tight schedule at FERMI, only a few shifts of several hours could be reserved to carry out these experiments. Since there was a positive outcome, it is evident, that the utilised methods are at an operational level.\\
	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
	\subsection{MFRL tests}
		\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
		\caption{The training of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_training}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
		\caption{The verification of the \emph{NAF2} on the FERMI FEL using a single network (dashed) and a double network (solid).}
		\label{fig:NAF_verification}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
		\caption{The training metrics of the \emph{AE-DYNA-SAC} on the FERMI FEL using a single network (dashed) and a double network (solid). The Bellman error (\cref{eq:minimize_bellmann_optimality}) and the state-value function (\cref{eq:state-value-function}) are shown. }
		\label{fig:NAF_convergence}
	\end{figure}
	In total, four tests were carried out, two using a single network and two using the double network architecture. A motivation for the used algorithmic set-up is discussed in \cref{appendix:naf2} on a classical example from control theory. In both cases \emph{smoothing} to increase the noise robustness as discussed in \cref{appendix:naf2}) was applied.\\
	\Cref{fig:NAF_training} displays the results, averaged over the two tests. A training of 100 episodes was accomplished. In the upper figure, the number of iterations per episode is plotted, including the cumulative number of steps. The green line in the lower plot of \cref{fig:NAF_training} indicates the 95\% intensity reward threshold, to accept an episode to be successfully finished.
	In the verification of 50 episodes (\cref{fig:NAF_verification}) both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance. \\
	Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps. The blue curves show the Bellmann error \cref{eq:minimize_bellmann_optimality}, which is comparable in both cases. The state-value function $V$ (\cref{eq:state-value-function}), which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value. In the single-network case, convergence is reached in ~1400 steps. $V$ was averaged over 200 states.
	\subsection{MBRL tests}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
		\caption{The training of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:AE-DYNA_observables}
	\end{figure}

@@

%	\section{Theoretical aspects on the \emph{AE-DYNA} on FERMI FEL}\label{appendix:Theoretical aspects on the 	AE-DYNA on FERMI FEL}
%	In thist section we want to discuss some theoretical aspects concerning the \emph{AE-DYNA} approach. A simulation was used to obtain the presented analysis. \\
%	One issue of MBRL methods is asymptotic performance. Although good results were obtained using the \emph{AE-DYNA}, there is a difference between the MFRL and MBRL. We call it the \emph{reality gap}. There are ideas to attack the problem, e.g. by learning a meta-policy \cite{Clavera2018} followed by a fine-tuning on the residual physics, which minimises the \emph{reality gap} of training on a simulator and closing than with a small number of training iterations on the real system the gap \cite{Zeng2019}.
%	\begin{figure}[!h]
%		\centering
%		\includegraphics*[width=0.5\textwidth]{Figures/Learning_evolution}
%		\caption{The training .}
%		\label{fig:Learning_evolution}
%	\end{figure}
\section{Still open}
	\begin{itemize}
	\item Reference clean up.
	\item Conclusion and outlook\dots
	\item Clean discussion of the results. What is superior and why? A table with numbers...
%	\item Noise benchmarks with pendulum - done
%	\item AE details from papers
%	\item Benchmark NAF2 with pendulum - done
%	\item Compare NAF2 and AE-DYNA
%	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
%	\item Tailored to accelerators - short horizons are standard?
%	\item Long waiting time in MBRL	
%	\item Vary the model size
%	\item Noise sensitivity as mentioned in Benchmark paper
%	\item Improvements of the algorithms:
%	\begin{itemize}
%		\item Adding the noise feature
%		\item Jumping feature using \emph{SAC}
%		\item Noise in the policy
%		\item discuss different approaches to, e.g. dynamic waiting time
%		\item usually prior not discussed in papers
%	\end{itemize}
\end{itemize}
